---
title: "Sectional Project 1"
author: "Group 3"
date: "2/11/2021"
output: pdf_document
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages('ggplot2')
library(ggplot2)
library(corrplot)

library(glmnet)
library(plotmo)

```

# Introduction to the Boston Housing Dataset


```{r importingData, echo=FALSE}

#Importing from directory; alternatively add the file to the github folder and load.
data <-read.table('https://archive.ics.uci.edu/ml/machine-learning-databases/housing/housing.data')

#Names (Note: The .names file contains metadata on attributes)
names(data) <-c("CRIM", "ZN", "INDUS", "CHAS", "NOX", "RM", "AGE", "DIS","RAD","TAX","PTRATIO","B","LSTAT","MEDV")

#There weren't any NA (dim(data) retained)
data <- na.omit(data)

```


The Boston Housing dataset considers housing values and their associated properties in suburbs of Boston, Massachusetts. The dataset contains 506 observations and 14 attributes. We acquired the dataset from the Machine Learning Database (MLDB), found [here](https://archive.ics.uci.edu/ml/machine-learning-databases/housing/). In particular, we are interested in constructing a model through regression techniques to gain insight on housing values. As such, we will use the 13 features to model 'MEDV', the median value of owner occupied homes (in $1,000s). 

The data is displayed as follows.
```{r displayData, echo=FALSE}
head(data)
```

# Exploratory Data Analysis


## MEDV Distribution

```{r MEDVHistogram, echo=FALSE,out.width='70%'}

histMEDV = ggplot(data, aes(MEDV)) + geom_histogram(bins=30,color="darkblue", fill="lightblue")+ ggtitle("Median Value Distribution") + theme(plot.title = element_text(hjust = 0.5, size = 17)) 

histMEDV
```

The histogram demonstrates the values are not uniformly distributed. Rather, they follow a mostly normal distributions, with some outliers at the tail.


## Correlation Matrix


```{r Corrmatrix, echo=FALSE,out.width='70%'}

corMatrix <- cor(data)
corrplot.mixed(corMatrix, number.cex= 9/ncol(data),tl.cex= 9/ncol(data))


```

While we produce many correlation values, we are firstly interested in how each attribute correlates to MEDV. This is represented by the bottom row or last column. We can immediately see the binary CHAS attribute does not correlate strongly with MEDV. However, it can be seen that RM (0.7) and LSTAT (-0.74) correlate with MEDV stronger than other attributes. Furthermore, the correlation between RM and LSTAT is -0.61. Since they do not correlate very strongly with one another, we can select both as predictor attributes without too much concern of collinearity for their case. The greatest correlation is between RAD and TAX of 0.91. Including both of these may raise some concerns regarding the minimal collinearity assumption of linear regression.


## Example EDA subsection title 3


# Modelling and Regression

MedV takes the value for Y, along 13 feature attributes of the dataset, in the form of $Y = \beta_0+\beta_1x_1+...+\beta_nx_n$.


## Multiple Linear Regression


This is very subject to change. I chose to simply split the data in a 80/20 split. If you think we should consider another split, or more, please do so. Furthermore, this is simply a start, and have yet to articulate everything. Please feel free to make changes.
```{r dataSplitting, echo=FALSE}
#Setting seed for reproducible results
set.seed(1)

indices <-sample(1:nrow(data), 0.6 * nrow(data), replace = TRUE)
training <-data[indices,]
testing  <-data[-indices,]
```

Naively consider most attributes at onset.
```{r multiModel1, echo=FALSE}
#Naively considering many attributes
multiModel1 <-lm(MEDV ~CRIM+ZN+INDUS+NOX+RM+AGE+DIS+RAD+TAX+PTRATIO+B+LSTAT, data=training)
summary(multiModel1)
```

Some initial insight is that LSTAT and RM indeed were strong predictors. Removing INDUS and AGE, every attribute becomes a significant predictor with the possible exception of B, depending on alpha. Let's consider what happens if we remove RAD, which varies strongly with TAX.

```{r multiModel2, echo=FALSE}
multiModel2 <-lm(MEDV ~CRIM+ZN+NOX+RM+DIS+TAX+PTRATIO+B+LSTAT, data=training)
summary(multiModel2)
```

We can see that without RAD, TAX is no longer a strong predictor. As such, TAX adds predictive value in relation to RAD. The next model removes TAX and adds RAD back in.

```{r multiModel3, echo=FALSE}
#Considering all values that are considered 'significant' with RAD included
multiModel3 <-lm(MEDV ~CRIM+ZN+NOX+RM+DIS+RAD+PTRATIO+B+LSTAT, data=training)
summary(multiModel3)
```

```{r multiModel4, echo=TRUE}
#This model is using any value that is abs(Intercept) > 1 for the ridge
multiModel4 <-lm(MEDV ~NOX+RM+DIS+PTRATIO+LSTAT, data=training)
summary(multiModel4)
```

MSE for model 1, 2, 3.
```{r multiModelTesting, echo=FALSE}
multiPredictions <-predict(multiModel1, testing)
RSS <- sum((testing$MEDV - multiPredictions)^2)
MSE1 <- mean((testing$MEDV - multiPredictions)^2)
MSE1

multiPredictions <-predict(multiModel2, testing)
RSS2 <- sum((testing$MEDV - multiPredictions)^2)
MSE2 <- mean((testing$MEDV - multiPredictions)^2)
MSE2

multiPredictions <-predict(multiModel3, testing)
RSS3 <- sum((testing$MEDV - multiPredictions)^2)
MSE3 <- mean((testing$MEDV - multiPredictions)^2)
MSE3
```

MSE for model 4, using only intercepts as derived from ridge coefficients.
```{r multiModelTesting2, echo=TRUE}
multiPredictions <-predict(multiModel4, testing)
MSE4 <- mean((testing$MEDV - multiPredictions)^2)
MSE4

# All Output are around 21 YET the ridge produces something like 42046. 
# How do we compare these two? Where is the error in my multiple lin regression?
```

Lasso with parameter tuning will give us further insight into paramater selection. This section can/should be expanded/refined.

## Ridge Regression

```{r ridgeDataFormatting, echo=FALSE}
#Setting seed
set.seed(1)

#Formatting into matrix
data.mat <- model.matrix(MEDV ~ .-1, data=data)

#delineating columns/features
x <- data.mat[, 1:13]
y <- data[, 'MEDV']

```

Constructing a ridge model.
```{r ridgeModelStart, echo=TRUE}
#lambda grid
grid <- 10^seq(6, -3, length=10)
#ridgeModel
ridge.mod <- glmnet(scale(x), y, alpha = 0, lambda = grid, thresh = 1e-2, standardize = TRUE)
plot_glmnet(ridge.mod, xvar = "lambda", label = 13)
```

Considering a best lambda for the model and hyperparameter tuning.
```{r ridgeLambdaTuning, echo=TRUE}
cv.out <- cv.glmnet(x, y, alpha=0, nfolds = 10)
cv.out

plot(cv.out)

best.lambda <- cv.out$lambda.min
best.lambda
```

Considering coefficients for ridge.
```{r ridgeCoeff, echo=TRUE}

ridge.final <- glmnet(scale(x), y, alpha = 0, lambda = best.lambda, thresh=1e-2, standardsize = TRUE)
predict(ridge.final, type="coefficients", s=best.lambda)

```

Considering MSE and RMSE for ridge using all coefficients.
```{r ridgeError, echo=FALSE}

ridge.pred <- predict(ridge.final, s=best.lambda, newx=x)
ridge.MSE = mean((ridge.pred -y)^2)
ridge.RMSE = sqrt(mean((ridge.pred - y)^2))

ridge.MSE
ridge.RMSE

```

## Lasso Regression

## K-Fold Cross Validation 

A simple quick table for us to store into for now:

|           | MSE | RMSE |
|-----------|-----|------|
| Multiple  | 0   | 0    |
| Multiple 2| 0   | 0    |
| Ridge     | 0   | 0    |
| Lasso     | 0   | 0    |



# Citations







