---
title: "Project 3"
author: "Group 3"
date: "4/4/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Group Leader: \underline{Brianna Johnson}

Member Names: \underline{Marco Sousa, Ben Pfeffer, Nikita Seleznev, Brianna Johnson}


# Introduction to the US Census Binary Income Dataset

```{r libraries, include=FALSE}
#Visualization
library(ggplot2)
#Utilities
library(dplyr)
#Regression
library(ISLR)
#correlations
library(corrplot)
# library to PLOT ROC
library(pROC)

```

Credited to Ronny Kohavi and Barry Becker, The  1994 United States Census Bureau data has several features describing an individual to predict in a binary manner whether that individual will earn more or less than $50,000 per year.

```{r importingData, echo=FALSE}
#Importing from directory. I specify a csv sep and strip whitespace
dataTrain <- read.table('https://archive.ics.uci.edu/ml/machine-learning-databases/adult/adult.data',sep=",",strip.white=TRUE)

#Names (Note: The .names file contains metadata on attributes)
names(dataTrain) <-c("age", "workclass", "fnlwgt", "education", "education.num", "marital.status", "occupation", "relationship","race","sex","capital.gain","capital.loss","hours.per.week","native.country","income")

# find elements of ? (similar to NA)
idx <- dataTrain == "?"
# replace elements with NA
is.na(dataTrain) <- idx

#Choosing to remove NA
dataTrain <- na.omit(dataTrain)

#Making binary column binary
dataTrain <- dataTrain %>%
      mutate(incomeBinary = ifelse(income == "<=50K",0,1))


dataTest <- read.table('adult.test',sep=",",strip.white=TRUE)

names(dataTest) <-c("age", "workclass", "fnlwgt", "education", "education.num", "marital.status", "occupation", "relationship","race","sex","capital.gain","capital.loss","hours.per.week","native.country","income")

# find elements
idx <- dataTest == "?"
# replace elements with NA
is.na(dataTest) <- idx

dataTest <- na.omit(dataTest)

dataTest <- dataTest %>%
  mutate(incomeBinary = ifelse(income == "<=50K.",0,1))

nrow(dataTrain)
nrow(dataTest)
```

The training data is displayed as follows.
```{r seeData, echo=FALSE}
head(dataTrain)
```


# Exploratory Data Analysis

## Binary Income Count

```{r incomeHistTrain, echo=FALSE,out.width='70%'}
incomeTrBar <- ggplot(dataTrain, aes(incomeBinary)) + geom_bar(fill="lightblue",color="black")+ ggtitle("Train Income <=50K vs >50K") + theme(plot.title = element_text(hjust = 0.5, size = 17)) + geom_text(stat='count', aes(label=..count..), vjust=5)

incomeTrBar
```

```{r incomeHistTest, echo=FALSE,out.width='70%'}
incomeTestBar <- ggplot(dataTest, aes(incomeBinary)) + geom_bar(fill="lightblue",color="black")+ ggtitle("Train Income <=50K vs >50K") + theme(plot.title = element_text(hjust = 0.5, size = 17)) + geom_text(stat='count', aes(label=..count..), vjust=5)

incomeTestBar
```

## Quick Histograms

```{r histograms, echo=FALSE,out.width='50%'}

trainAge = hist(dataTrain$age)
trainFnl = hist(dataTrain$fnlwgt)
trainGain = hist(dataTrain$capital.gain)
trainLoss = hist(dataTrain$capital.loss)
```

## EDA Next


# Logistic Regression




# Naive Bayes



