---
title: "Project 2 - Regression and CART"
author: "Group 3"
date: "3/13/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Group Leader: \underline{Brianna Johnson}

Member Names: \underline{Marco Sousa, Ben Pfeffer, Nikita Seleznev, Brianna Johnson}


# Introduction to the Heart Dataset

```{r libraries, include=FALSE}
library(ggplot2)

library(dplyr)

```

Observing the data as its initial dataframe.
```{r loadingData, echo=FALSE}

#Importing from directory
data <- read.csv("https://www.statlearning.com/s/Heart.csv ")

#Dropping first bogus column
data = data[-c(1)]

#Constructs a new column replacing No with 0 and Yes with 1
data <- data %>%
      mutate(AHDBinary = ifelse(AHD == "No",0,1))

head(data)
```

Describing the data, and (1) identify Y and X.

I imagine the feature we are interested in is AHD, which is either "Yes" or "No" in regards to having heart disease. X are the other possible features which may predict Y.


# Exploratory Data Analysis

EDA

# Logistic Model

## Logistic Model

Perhaps construct logistic model.

## Confusion Table

Generate confusion table.
 
## Mode-Test Statistics

Calculate classification accuracy and error, sensitivity, specificity, PPV and NPV. Will need to be compared elsewhere

## ROC and AUC

Generate ROC and compute AUC for each model

## S sigmoid curve

Generate s-curve for Y against one attribute (you can pick any one attribute), and interpret your findings


# CART Model

## CART MODEL

## Other subsections

# Comparing Models

# Citations

